Learning and comparing functional connectomes across subjects

### 2013-01-04 paper.tex
Small changes

Predictive modeling is concerned with learning (or fitting) a model that is
capable of decoding or \emph{predicting} information from new unseen data
\cite{pereira2009}.  In the context of connectomes, predictive modeling provides
a means for delineating connectivity based biomarkers of disease diagnosis,
prognosis, or other phenotypic outcomes \cite{craddock2009, dosenbach2010}. The
accuracy of a predictive model provides a measure of the amount of information
present in the connectome about the phenotypic measure being evaluated
\cite{kjems2002, kriegeskorte2006}.  When combined with reproducibility, prediction accuracy
provides a metric for evaluating experimental tradeoffs for data acquisition,
preprocessing, and analysis \cite{strother2002, laconte2003}.  Multivariate
predictive modeling (MPM)
are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisons since
the significance of an entire pattern is evaluated using a single statistical
test. 
Additionally many modern MPM techniques exist which are specifically
formulated to deal with high dimensional datasets with few observations.
Predictive modeling techniques have been
successfully applied to identify connectome-based biomarkers of Alzheimer's
disease \cite{stonnington2010}, depression \cite{craddock2009, zeng2012}, schizophrenia
\cite{cecchi2009, shen2010}, autism \cite{anderson2011}, ADHD \cite{zhu2008}, aging
\cite{dosenbach2010}, as well as to classify mental operations
\cite{richiardi2011,shirer2012}. The growing interest for applying MPM 
to connectivity analysis was highlighted by the ADHD200
Global Competition, in which the object was to identify a connectivity based
biomarker of ADHD \cite{adhd2002012}. Recent work has illustrated the utility
of predictive modeling for deriving connectivity models at the individual level
\cite{chu2011}.

@@

% A bit of technical details
% technical details
Technically, predictive modeling is a supervised machine learning problem where
a target to be predicted --\emph{e.g.}\ age, disease state, cognitive state-- is
available for each observation of the data. In the context of comparing
connectomes, features used in the predictive model correspond to bivariate
measures of connectivity, or any of the previously discussed summary metrics.  
%
The quality of a predictive model is determined by its prediction accuracy (or
generalization ability) which is measured using one or more iterations of
cross-validation. Cross-validation iteratively seperates available data into a
subset used for training the classifier and a dataset for evaluating classifier
performance. Several strategies exist for performing cross-valdiaiton and the commonly
used approach of using only a single observation for testing (leave-one-out
cross-validation) results in highly variable estimates of prediction accuracy
\cite{hastie2003}.
Alternative approaches such as (5 or 10)-fold cross-validation, split half
resampling, or $0.632+$ bootstrap should be preferred \cite{hastie2003}. The significance of
achieved prediction accuracy can be assessed using permutation tests
\cite{golland2003}.
Predictive modeling approaches typically require the specification of several
parameters, which may be chosen based on domain specific knowledge or
requirements \cite{cherkassky1998}, determined using an analytical approach
\cite{cherkassky2004}, or optimized using a second level cross validation procedure
\cite{cherkassky1998}.

Although MPM techniques are extremely well suited at measuring whether
information exists in the connectome about a phenotypic variable, indentifying
the connections most relevant to the predictive model (feature selection) is
less straight forward.  These problems can be somewhat mitigated by carefully
previously described sparse inverse covariance estimation techniques to minimize
the number of connections. Feature selection \cite{guyon2003} can be performed by filtering out
features based on their statistical relationship with the variable of interest
\cite{craddock2009, shen2010},
selecting features based on cross-validation \cite{craddock2009, deshpande2010}, or by employing sparse modeling
methods that force weights for non-essential features to be zero
\cite{tibshirani1994, michel2009, ryali2010}.
Filtering features based on a univariate criterion such as t-test should be
avoided, since they are insensitive to multivariate information and may exclude
features that could be important to a multivariate model \cite{guyon2003,
	craddock2009}. No matter which
feature selection strategy is employed, it must be performed within
cross-validation to avoid biasing estimates of prediction accuracy.

>>>>

Predictive modeling is concerned with learning (or fitting) a model that is
capable of \emph{predicting} information from new unseen data
\cite{pereira2009}.  In the context of connectomes, predictive modeling provides
a means for delineating connectivity based biomarkers of disease diagnosis,
prognosis, or other phenotypic outcomes \cite{craddock2009, dosenbach2010}. The
accuracy of a predictive model provides a measure of the amount of information
present in the connectome about the phenotypic measure being evaluated
\cite{kjems2002, kriegeskorte2006}.  When combined with reproducibility, prediction accuracy
provides a metric for evaluating experimental tradeoffs for data acquisition,
preprocessing, and analysis \cite{strother2002, laconte2003}.  Multivariate
predictive models are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisons since
the significance of an entire pattern is evaluated using a single statistical
test. 
Additionally modern predictive modeling techniques draw from the
statistical learning literature, which specifically addresses
 high dimensional datasets with few observations.
Predictive modeling has been
successfully applied to identify connectome-based biomarkers of Alzheimer's
disease \cite{stonnington2010}, depression \cite{craddock2009, zeng2012}, schizophrenia
\cite{cecchi2009, shen2010}, autism \cite{anderson2011}, ADHD \cite{zhu2008}, aging
\cite{dosenbach2010}, as well as to classify mental operations
\cite{richiardi2011,shirer2012}. The growing interest for applying
predictive modeling to connectivity analysis was highlighted by the ADHD200
Global Competition, in which the object was to identify a connectivity based
biomarker of ADHD \cite{adhd2002012}. Recent work has illustrated the utility
of predictive modeling for deriving connectivity models at the individual level
\cite{chu2011}.

@@

% A bit of technical details
% technical details
Technically, predictive modeling is a supervised machine learning problem where
a target to be predicted --\emph{e.g.}\ age, disease state, cognitive state-- is
available for each observation of the data. In the context of comparing
connectomes, features used in the predictive model correspond to bivariate
measures of connectivity, or any of the previously discussed summary metrics.  
%
The quality of a predictive model is determined by its prediction accuracy (or
generalization ability) which is measured using one or more iterations of
cross-validation. Cross-validation iteratively separates available data into a
subset used for training the classifier and a dataset for evaluating classifier
performance. Several strategies exist for performing cross-validation and the commonly
used approach of using only a single observation for testing (leave-one-out
cross-validation) results in highly variable estimates of prediction accuracy
\cite{hastie2003}.
Alternative approaches such as (5 or 10)-fold cross-validation, split half
resampling, or $0.632+$ bootstrap should be preferred \cite{hastie2003}. The significance of
achieved prediction accuracy can be assessed using permutation tests
\cite{golland2003}.
Predictive modeling approaches typically require the specification of several
parameters, which may be chosen based on domain specific knowledge or
requirements \cite{cherkassky1998}, determined using an analytical approach
\cite{cherkassky2004}, or optimized using a second level cross validation procedure
\cite{cherkassky1998}.

Although predictive modeling techniques are well suited at measuring whether
information exists in the connectome about a phenotypic variable, they do
not directly identify the connections that are most relevant to the prediction.  These problems can be somewhat mitigated  by
relying on previously described sparse inverse covariance estimation techniques to minimize
the number of connections. In addition, feature selection \cite{guyon2003} 
can be performed by filtering out
features based on their statistical relationship with the variable of interest
\cite{craddock2009, shen2010},
selecting features based on cross-validation \cite{craddock2009, deshpande2010}, or by employing sparse modeling
methods that force weights for non-essential features to be zero
\cite{tibshirani1994, michel2009, ryali2010}.
Filtering features based on a univariate criterion such as t-test should be
avoided, since they are insensitive to multivariate information and may exclude
features that could be important to a multivariate model \cite{guyon2003,
	craddock2009}. No matter which
feature selection strategy is employed, it must be performed within
cross-validation to avoid biasing estimates of prediction accuracy.
